---
title: Coursera Statistical Inference Project - Part 2
author: Brian Altman
date: December 26, 2015
output: pdf_document
---

##**Overview**      
The project analyzes the ToothGrowth data in the R datasets package. The project 

      1) Load the ToothGrowth data
      2) Perform some basic exploratory data analyses and a basic summary of the data.
      3) Uses confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
      4) Provides conclusions and the assumptions needed

This file and related code can be found here: https://github.com/beaisis/Statistical-Inference 
***

####**1) Load the ToothGrowth data**
Tooth growth data is part of the base R package install.
```{r data}
library(ggplot2)
data(ToothGrowth)
```

####**2) Perform some basic exploratory data analyses related to supp and dose**
There are 60 observations of  3 variables:
```{r Basic}

str(ToothGrowth)
head(ToothGrowth)

summary(ToothGrowth)
```
Means of tooth grown by dose and supp(delivery method)
```{r Basic2}
tapply(ToothGrowth$len, list(ToothGrowth$dose,ToothGrowth$supp),mean)
```
Plot showing tooth growth by dose and supp:

```{r Basic3, echo=FALSE}
ggplot(data=ToothGrowth, aes(x=ToothGrowth$dose, y=len, fill=supp)) + geom_boxplot() +
      facet_grid(. ~ supp) +
      xlab("Dose (Milligrams)") + ylab("Tooth Length") + ggtitle("Tooth Growth by Supp and Dose")
```
The summary tables and boxplot suggests tooth length is positively correlated with the dosage for both supplements.

####**3) Confidence intervals and null hypothesis tests to compare tooth growth by supp and dose**



*Null hypothesis: No correlation between method of delivery and tooth length.*

The following t-test was conducted:
```{r T Test}
t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"], ToothGrowth$len[ToothGrowth$supp=="VC"], paired = FALSE, var.equal = FALSE)
```
A P-value of .06 and confidence interval of [-0.171, 7.571] that includes zero, suggests the null hypthesis can not be rejected. There does not appear to be a correlation between method of delivery and tooth length.



*Null Hypothesis: No correlation between dosage and Tooth Length*. 
The following models by dose and supp were run:

```{r Dosage, echo=FALSE}
t_tests <- ""
dose <-as.factor(ToothGrowth$dose)
len  <-ToothGrowth$len
supp <-ToothGrowth$supp
```

```{r Models}
t1<-t.test(len[dose==.5 & supp == 'VC'],len[dose==1& supp == 'VC'], paired = FALSE, var.equal = TRUE)
t_tests <- rbind(t_tests, list('.5~1 mg - VC',  t1$p.value,t1$conf.int[1], t1$conf.int[2]))
t1<-t.test(len[dose==1 & supp == 'VC'], len[dose==2 & supp == 'VC' ], paired = FALSE, var.equal = TRUE)
t_tests <- rbind(t_tests, list(' 1~2 mg - VC',  t1$p.value, t1$conf.int[1], t1$conf.int[2]))
t1<-t.test(len[dose==.5 & supp == 'VC'], len[dose==2 & supp == 'VC'], paired = FALSE, var.equal = TRUE)
t_tests <- rbind(t_tests, list('.5~2 mg - VC',  t1$p.value,t1$conf.int[1], t1$conf.int[2]))

t1<-t.test(len[dose==.5 & supp == 'OJ'],len[dose==1 & supp == 'OJ'], paired = FALSE, var.equal = TRUE)
t_tests <- rbind(t_tests, list('.5~1 mg - OJ',  t1$p.value,t1$conf.int[1], t1$conf.int[2]))
t1<-t.test(len[dose==1 & supp == 'OJ'], len[dose==2 & supp == 'OJ' ], paired = FALSE, var.equal = TRUE)
t_tests <- rbind(t_tests, list(' 1~2 mg - OJ',  t1$p.value,t1$conf.int[1], t1$conf.int[2]))
t1<-t.test(len[dose==.5 & supp == 'OJ'], len[dose==2 & supp == 'OJ'], paired = FALSE, var.equal = TRUE)
t_tests <- rbind(t_tests, list('.5~2 mg - OJ',  t1$p.value,t1$conf.int[1], t1$conf.int[2]))
colnames(t_tests) = c('Model','p.value','Ci.lower','ci.upper')
t_tests <- t_tests[2:6,]
```

Summary of the 6 models that represented the 3 Dose and 2 delivery method combinations:

```{r T Summaries, echo=FALSE}
t_tests
```

For all dosage combinations, the p-values are less than .05 and the confidence intervals do not contain zero.The null hypothesis is rejected: Dosages of .5, 1 and 2 do contribute to tooth length increases. 

####**4)Conclusions** 

Assuming random assignment of guinea pigs to different dose level categories and supplement types:

      Supplement type has no effect on tooth growth.

      Increasing the dose level leads to increased tooth growth.




