---
title: "Coursera Statistical Inference Project - Part 1"
author: "Brian Altman"
date: "December 25, 2015"
output: pdf_document
---

##**Overview**      
This project explores R's exponential distribution and compares it with the Central Limit Theorem. The exponential distribution is simulated in R with rexp(n, lambda). The project's objectives:

 1. Create a simulation using rexp(n, lambda)
 2. Compare the simulation mean to the theoretical mean of the distribution.
 3. Compare the simulation variance to the theoretical variance of the distribution
 4. Show the sample distribution is approximately normal.
 
This file and related code can be found here: https://github.com/beaisis/Statistical-Inference

####**Simulation**
R's rexp(n, lambda), where lamda is the rate parameter, was used to create a matix of 1000x40 random exponential numbers. With simulation parameters given by the project requirements, the means, variance and confidence intervals are determined for both the simulation and the theoretical models.

```{r Simulation}
library(ggplot2)
set.seed(314)
simulations   <- 1000  #Given by the project requirements.
lambda        <- 0.2   #Given by the project requirements.
n             <- 40    #Given by the project requirements.

simulated_means <- apply(matrix(rexp(simulations*n, lambda), simulations, n), 1, mean)

simulated_mean <- mean(simulated_means)
simulated_var  <- var(simulated_means)
simulated_std  <- sd(simulated_means)
simulated_ci   <- simulated_mean + c(-1,1)*1.96*sd(simulated_means)/sqrt(n)

theoretical_mean <- 1/lambda   #Given by the project requirements.
theoretical_var  <- 1/(lambda^2*n)
theoretical_std  <- (1/lambda)/sqrt(n)
theoretical_ci   <- theoretical_mean + c(-1,1)*1.96*theoretical_std/n
```


####**Compare the sample mean to the theoretical mean of the distribution.**

The following figure shows the density for the simulation, the mean of the simulation (dashed blue line) and the theoretical mean (yellow).


```{r Figure 1, echo=FALSE}
simulated_means_df <- data.frame(simulated_means)
g <- ggplot(simulated_means_df, aes(x = simulated_means)) 
g <- g + geom_histogram(aes(y=..density..), 
                        colour="blue", 
                        fill = "grey", binwidth=.1) 
g <- g + xlab("Simulated Means") + ylab("Density") + geom_density(colour="black", size=1) 
g <- g + ggtitle("Figure 1: Distribution of Simulated means")
g <- g + geom_vline(xintercept = simulated_mean, colour="blue", linetype = "longdash")
g <- g + geom_vline(xintercept = theoretical_mean, colour="yellow" )
g
```
Refering to above Figure 1, the simulation and theoretical mean are visually the same. The values indicate less than 1% difference:

      Simulated mean  : `r simulated_mean`
      Theoretical mean: `r theoretical_mean`


####**Compare the simulated and theoretical variance.**
The simulated and theoretical standard deviation and variance are as follows:

      Standard Deviation
            Simulated std  : `r simulated_std`
            Theoretical std: `r theoretical_std`
      Variance
            Simulated var  : `r simulated_var`
            Theoretical var: `r theoretical_var`


####**Confirm the sample distribution is approximately normal** 

In addition to Figure 1, the following shows the distribution compared to the normal distribution and evaluates the coverage of the confidence interval for 1/lambda: X¯±1.96S/n.

```{r QQ}
qqnorm(simulated_means)
qqline(simulated_means, col = 2)
```
The confidence intervals for the simulation and the theoretical are:

      Simulated CI  : `r simulated_ci`
      Theoretical CI: `r theoretical_ci`

####**Conclusion**

The distribution is very close to a normal distribution, supporting the Central Limit Theorem
